---
title: "Simple linear regression on noisy dataset with known parameters"
author: "Nikita Barsukov"
date: "15 dec 2017"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(reshape2)
library(scales)
library(caTools)
library(knitr)
options(scipen=999)

salary = function(experience) {
  10000 * experience + 15500
}
R.squared = function(actual, predicted) {
  1 - (sum((actual - predicted) ^ 2) / sum((actual - mean(actual)) ^ 2))
}

set.seed(2017)

```

# Linear regression on a noisy dataset with known parameters.

----------------
This is a first text in a series of writings for an online course "Machine Learning A-Z™: Hands-On Python & R In Data Science".

----------------

Most of the tutorials on linear regression 

What is interesting is to see how this simple algorithm performes on a data with *known* parameters and with added random noise.

I came up with this base equation that shows relationship between salary and years in job market: 

$Salary = 10.000 * experience + 15.500$

Then I generated 2.000 numbers for independent variable:

 * uniform distribution, 
 * between 0 and 20 inclusive,
 * rounded to nearest decimal place.

```{r}
observations = 2000
experiences = round(runif(observations, 0, 20), digits = 1)
```

And finally I constructed a data frame with freshly generated years of experience, salary calculated using our base formula and noisy salary: base-formula salary with added random noise.

To calculate noise I multiplied two uniformly distributed variables. Both had mean 0, one had a large standard deviation (thousands). Another has single-digit standard deviation. Salary with added noise is taken module if it is less than 0 and rounded to nearest hundred.

```{r}
dataset = data.frame(
  years_of_exp = experiences,
  salaries_true = salary(experiences),
  salaries_observed = abs(
    salary(experiences) + round(rnorm(observations, sd = 10000) * rnorm(observations, sd = 5), -2)
  )
)

```

First 10 rows of our generated dataset looks like this:

```{r echo=FALSE, results='asis'}
kable(dataset[1:10,], caption = "Sample of dataset", format.args = list(decimal.mark = ",", big.mark = "."))

```

Let's plot our noisy dataset.

```{r dataset, echo=FALSE}
pl = ggplot(data = dataset, aes(x=years_of_exp, y=salaries_observed)) +
  geom_point(color="#855C75") + 
  scale_x_continuous(name = "Years after graduation") +
  scale_y_continuous(name = "Observed salary", labels = comma) +
  ggtitle("Linear regression on noisy data from a base line", subtitle = "Random dataset generated from base line") +
  theme_minimal()

print(pl)
```

We see a clear linear pattern albeit with a substaitial noise around base axis.

Our next step is to split our dataset into training a test sets. This is a standard machine learning technique.
